full title

### 2020-11-03 paper/pldi21/javascript.tex
Added labels for each section
Added running example / Citation for TOPLAS'18
Changed: lazy concrete execution -> sealed symbolic execution
Revised Section 3,4

\section{Application: JavaScript}

In this section, we introduce the core language of JavaScript that supports
first-class functions, open objects, and first-class property names, and define
combined analysis of the core language.

@@

A program $\prog$ is a sequence of labelled instructions. An instruction $\inst$
is an expression assignment, an object creation, a function call, a return
instruction, or a branch.  A reference $\refer$ is a variable or a property
access of an object.  An expression $\expr$ is a primitive, a lambda function, a
reference, or an operation between other expressions.

@@

We formulate the concrete semantics of the core language as described in
Figure~\ref{fig:core-trans-rel}.  The transition relation between concrete
states is defined with the semantics of references and expressions using two
different forms \fbox{$\referrule{\st}{\refer}{\loc}$} and
\fbox{$\exprrule{\st}{\expr}{\val}$}, respectively.  The initial states are
$\istset = \{ (\ilab, \varnothing, \epsilon, \tladdr) \}$ where $\ilab$ denotes
the initial label, $\epsilon$ empty map, and $\tladdr$ the top-level
environment address.  The function $\labnext$ returns the next lable of a
given lable in the current program $\prog$.

@@

\subsection{Lazy Concrete Interpretation}

>>>>

\section{Dynamic Shortcut for JavaScript}\label{sec:javascript}

In this section, we introduce the core language of JavaScript that supports
first-class functions, open objects, and first-class property names, and define
sealed symbolic execution of the core language for dynamic shortcut.

@@

A program $\prog$ is a sequence of labeled instructions. An instruction $\inst$
is an expression assignment, an object creation, a function call, a return
instruction, or a branch.  A reference $\refer$ is a variable or a property
access of an object.  An expression $\expr$ is a primitive, a lambda function, a
reference, or an operation between other expressions.

@@

We formulate the concrete semantics of the core language as described in
Figure~\ref{fig:core-trans-rel}.  The transition relation between concrete
states is defined with the semantics of references and expressions using two
different forms \fbox{$\referrule{\st}{\refer}{\loc}$} and
\fbox{$\exprrule{\st}{\expr}{\val}$}, respectively.  The initial states are
$\istset = \{ (\ilab, \varnothing, \epsilon, \tladdr) \}$ where $\ilab$ denotes
the initial label, $\epsilon$ empty map, and $\tladdr$ the top-level
environment address.  The function $\labnext$ returns the next label of a
given label in the current program $\prog$.

@@

\subsection{Sealed Symbolic Execution}

### 2020-11-03 paper/pldi21/shortcut.tex
Revised Section 3,4

\section{Combined Analysis}\label{sec:formal}

In this section, we formally define the combined analysis with sealed symbolic
execution.  We extend the formalization of abstract interpretation and analysis
sensitivity proposed by \citet{sens-toplas}.

@@

For example, the code in Figure~\ref{fig:running-example} is a simple program
that applies the mathematical absolute value function to the variable $\varx$.
The question mark $?$ denotes the user input that returns an integer.  States
are pairs of labels and integers $\varx$: $\stset = \labset \times \mathbb{N}$.
The initial states are $\istset = \{ (\lab_0, 0) \}$ which means the program
starts with the varible $\varx$ that stores 0 at $\lab_0$.  If the user input is
$-42$, the program is executed with the following trace:
\[
  (\lab_0, 0) \trans (\lab_1, -42) \trans (\lab_3, -42) \trans (\lab_4, 42)
\]

@@

With analysis sensitivities, the abstract one-step execution $\sabsstep:
\sabsdom \rightarrow \sabsdom$ is defined as follows:
\[
  \sabsstep(\sabselem) = \lambda \view \in \viewset. \; \underset{\view' \in
  \viewset}{\bigjoin}{\viewtrans{\view'}{\view} \circ \sabselem(\view')}
\]
where $\viewtrans{\view'}{\view}: \absdom \rightarrow \absdom$ is the abstract
semantics of a \textit{view transition} from a view $\view'$ to another view
$\view$.  It should satsify the following condition for the soundness of the
analysis:
\[
  \forall \abselem \in \absdom. \; \step(\gamma(\abselem) \cap \viewmap(\view'))
  \cap \viewmap(\view) \subseteq \gamma \circ
  \viewtrans{\view'}{\view}(\abselem)
\]

@@

The main difference of sealed symbolic execution with the traditional symbolic
execution~\cite{symbolic} is that it only supports sealed symbolic values
instead of symbolic expressions and path constraints.  For example, the
following trace represents the traditional symbolic execution of the running
example in Figure~\ref{fig:running-example}:
\[
  \begin{array}{r@{~}c@{~}l}
    &&(\lab_2, \symb)[\symb \geq 0] \trans (\lab_4, \symb)[\symb \geq 0]
    \vspace*{-0.5em}\\
    &\rutrans&
    \vspace*{-0.5em}\\
    (\lab_0, 0)[] \trans (\lab_1, \symb)[]&&
    \vspace*{-0.5em}\\
    &\rdtrans&
    \vspace*{-0.5em}\\
    &&(\lab_3, \symb)[\symb < 0] \trans (\lab_4, -\symb)[\symb < 0]\\
  \end{array}
\]
It first assigns a symbolic value $\symb$ to the variable $\varx$ at $\lab_1$.
For the conditional branch, it is forked by creating two symbolic states with
different path conditions $\symb \geq 0$ and $\symb < 0$ for true and false
branches, respectively.  After executing statements $\varx = \varx$ and $\varx =
-\varx$, the variable $\varx$ stores symblic expressions $\symb$ and $-\symb$ at
$\lab_4$.  However, the sealed symbolic execution stops at $\lab_1$ as follows:
\[
  (\lab_0, 0) \; \symbtrans \; (\lab_1, \symb) \; \symbtrans \; \excst
\]
because the branch requires the actual value of the symbolic value $\symb$.

To freely convert a pair of view and an abstract state to its corresponding
symbolic state and vice versa, we define two domain converters $\symbstset
\galois{\saconverter}{\asconverter} (\viewset \times \absdom)$.  In the
converter $\asconverter$ from abstract states to symbolic states, if an abstract
value $\absval$ represents a singleton value, the function transforms the
abstract value to the corresponding concrete value.  Otherwise, it converts the
abstract value as a symbolic value in a result symbolic state.  Two converters
convert given elements wihtout loss of information:
\[
  \saconverter \circ \asconverter = \asconverter \circ \saconverter = \identity
\]
where $\identity$ denotes the identity function.

\subsection{Combined Analysis}

\begin{figure}[t]
  \centering
  \includegraphics[width=\linewidth]{img/combined-analysis}
  \vspace*{-2em}
  \caption{The diagram for the extended view transition in the combined analysis.}
  \vspace*{-1em}
  \label{fig:combined-analysis}
\end{figure}

We define the \textit{combined analysis} of the abstract interpretation and the
sealed symbolic execution.  In the combined analysis, we use the powerset of
symbolic states as the domain $\combdom = \powerset{\symbstset}$, and extend the
view transition as follows:
\[
  \begin{array}{l}
    \combviewtrans{\view}{\view'}(\combelem) =\\
    \qquad \{
      \symbst' \mid \symbst \in S \wedge
      \symbst \symbtrans \symbst' \wedge
      \saconverter(\symbst') = (\view', \_)
    \}\\
    \qquad \cup \left\{
    \begin{array}{ll}
      \{ \asconverter \circ \viewtrans{\view}{\view'}(\bigjoin D) \}
      & \text{if} \; D \neq \varnothing\\
      \varnothing & \text{otherwise}\\
    \end{array}
    \right.
  \end{array}
\]
where $S = \combelem\mid_{\checker}$ and $D =
\dot{\saconverter}(\combelem\mid_{\neg\checker})$.  The dot notation $\dot{f}$
denotes the element-wise extended function of a function $f$.  The combined
analysis performs one-step symbolic transition for each symbolic state that
passes the filter $\checker$. On the other hand, it converts remaining symbolic
states to the corresponding abstract states, merges them into a single abstract
state, and performs the abstract one-step execution for the abstract state.

For example, Figure~\ref{fig:combined-analysis} depicts the
diagram of the view transition from $\view$ to $\view'$ in the combined
analysis.  The element $\combelem$ in $\view$ consists of five symbolic states
from $\nsymbst{1}$ to $\nsymbst{5}$.  The solid box denotes that the corresponding
symbolic state passes the filter $\checker$ and the dotted box denotes that it
does not.  For each symbolic state that passes the filter ($\nsymbst{1}$,
$\nsymbst{2}$, and $\nsymbst{4}$), it applies the symbolic transition $\symbtrans$.
For failed symbolic states ($\nsymbst{3}$ and $\nsymbst{5}$), it applies the
converter $\saconverter$ to convert each of them to the corresponding abstract
state and merges results using the join operator ($\join$).  Then, it utilizes
the abstract semantics via the original view transition
$\viewtrans{\view}{\view'}$ and the converter $\asconverter$ to convert the
result to the corresponding symbolic state.

We could configure when the sealed symbolic execution or the abstract
interpretation is performed based on the definition of the $\checker$ function
and its negation $\neg\checker$.  For example, we could define the $\checker$
that passes only function entry, call, and exit points to perform conversions in
a function level.  However, for the soundness and the termination of the
combined analysis, it should satisfy the following condition:
\begin{theorem}
  The combined analysis has sound analysis result and terminates in a finite
  time if its abstract semantics is sound and $\checker$ satisfies the following
  condition:
  \[
    \checker(\symbst) \Rightarrow \symbst \symbtrans^k \excst \wedge
    1 < k \leq N
  \]
  where $N$ is a pre-defined maximum length of the sealed symbolic execution.
\end{theorem}
\inred{We prove that any combined analysis defined with the extended view
transition satisfying the above condition always has sound analysis result and
terminates in a finite time. Because of the page limitation, we omit the
proof in this paper and include it in a companion report.}

>>>>

\section{Dynamic Shortcut}\label{sec:formal}

In this section, we formally define the dynamic shortcut with sealed symbolic
execution.  We extend the formalization of abstract interpretation of
\citet{abs-interp-1977, abs-interp-1992} and views-based analysis sensitivity of
\citet{sens-toplas}.

@@

For example, the code in Figure~\ref{fig:running-example} is a simple program
that applies the mathematical absolute value function to the variable $\varx$.
The question mark $?$ denotes the user input that returns an integer.  States
are pairs of labels and integers $\varx$: $\stset = \labset \times \mathbb{N}$.
The initial states are $\istset = \{ (\lab_0, 0) \}$ which means the program
starts with the variable $\varx$ that stores 0 at $\lab_0$.  If the user input is
$-42$, the program is executed with the following trace:
\[
  (\lab_0, 0) \trans (\lab_1, -42) \trans (\lab_3, -42) \trans (\lab_4, 42)
\]

@@

With analysis sensitivities, the abstract one-step execution $\sabsstep:
\sabsdom \rightarrow \sabsdom$ is defined as follows:
\[
  \sabsstep(\sabselem) = \lambda \view \in \viewset. \; \underset{\view' \in
  \viewset}{\bigjoin}{\viewtrans{\view'}{\view} \circ \sabselem(\view')}
\]
where $\viewtrans{\view'}{\view}: \absdom \rightarrow \absdom$ is the abstract
semantics of a \textit{view transition} from a view $\view'$ to another view
$\view$.  It should satisfy the following condition for the soundness of the
analysis:
\[
  \forall \abselem \in \absdom. \; \step(\gamma(\abselem) \cap \viewmap(\view'))
  \cap \viewmap(\view) \subseteq \gamma \circ
  \viewtrans{\view'}{\view}(\abselem)
\]

@@

The main difference of sealed symbolic execution with the traditional symbolic
execution~\cite{symbolic} is that it only supports sealed symbolic values
instead of symbolic expressions and path constraints.  For example, the
following trace represents the traditional symbolic execution of the running
example in Figure~\ref{fig:running-example}:
\[
  \begin{array}{r@{~}c@{~}l}
    &&(\lab_2, \symb)[\symb \geq 0] \trans (\lab_4, \symb)[\symb \geq 0]
    \vspace*{-0.5em}\\
    &\rutrans&
    \vspace*{-0.5em}\\
    (\lab_0, 0)[] \trans (\lab_1, \symb)[]&&
    \vspace*{-0.5em}\\
    &\rdtrans&
    \vspace*{-0.5em}\\
    &&(\lab_3, \symb)[\symb < 0] \trans (\lab_4, -\symb)[\symb < 0]\\
  \end{array}
\]
It first assigns a symbolic value $\symb$ to the variable $\varx$ at $\lab_1$.
For the conditional branch, it is forked by creating two symbolic states with
different path conditions $\symb \geq 0$ and $\symb < 0$ for true and false
branches, respectively.  After executing statements $\varx = \varx$ and $\varx =
-\varx$, the variable $\varx$ stores symbolic expressions $\symb$ and $-\symb$ at
$\lab_4$.  However, the sealed symbolic execution stops at $\lab_1$ as follows:
\[
  (\lab_0, 0) \; \symbtrans \; (\lab_1, \symb) \; \symbtrans \; \excst
\]
because the branch requires the actual value of the symbolic value $\symb$.

To freely convert a pair of view and an abstract state to its corresponding
symbolic state and vice versa, we define two domain converters $\symbstset
\galois{\saconverter}{\asconverter} (\viewset \times \absdom)$.  In the
converter $\asconverter$ from abstract states to symbolic states, if an abstract
value $\absval$ represents a singleton value, the function transforms the
abstract value to the corresponding concrete value.  Otherwise, it converts the
abstract value as a symbolic value in a result symbolic state.  Two converters
convert given elements without loss of information:
\[
  \saconverter \circ \asconverter = \asconverter \circ \saconverter = \identity
\]
where $\identity$ denotes the identity function.

\subsection{Dynamic Shortcut}

\begin{figure}[t]
  \centering
  \includegraphics[width=\linewidth]{img/dynamic-shortcut}
  \vspace*{-2em}
  \caption{The diagram for the extended view transition for dynamic shortcut.}
  \vspace*{-1em}
  \label{fig:dynamic-shortcut}
\end{figure}

We define the the abstract interpretation with \textit{dynamic shortcut}, which
is a concrete semantics on sealed symbolic execution.  With dynamic shortcut,
we use the powerset of symbolic states as the domain $\combdom =
\powerset{\symbstset}$, and extend the view transition as follows:
\[
  \begin{array}{l}
    \combviewtrans{\view}{\view'}(\combelem) =\\
    \qquad \{
      \symbst' \mid \symbst \in S \wedge
      \symbst \symbtrans \symbst' \wedge
      \saconverter(\symbst') = (\view', \_)
    \}\\
    \qquad \cup \left\{
    \begin{array}{ll}
      \{ \asconverter \circ \viewtrans{\view}{\view'}(\bigjoin D) \}
      & \text{if} \; D \neq \varnothing\\
      \varnothing & \text{otherwise}\\
    \end{array}
    \right.
  \end{array}
\]
where $S = \combelem\mid_{\checker}$ and $D =
\dot{\saconverter}(\combelem\mid_{\neg\checker})$.  The dot notation $\dot{f}$
denotes the element-wise extended function of a function $f$.  The abstract
interpretation with dynamic shortcut performs one-step symbolic transition for
each symbolic state that passes the filter $\checker$. On the other hand, it
converts remaining symbolic states to the corresponding abstract states, merges
them into a single abstract state, and performs the abstract one-step execution
for the abstract state.

For example, Figure~\ref{fig:dynamic-shortcut} depicts the
diagram of the extended view transition from $\view$ to $\view'$ for dynamic
shortcut.  The element $\combelem$ in $\view$ consists of five symbolic states
from $\nsymbst{1}$ to $\nsymbst{5}$.  The solid box denotes that the corresponding
symbolic state passes the filter $\checker$ and the dotted box denotes that it
does not.  For each symbolic state that passes the filter ($\nsymbst{1}$,
$\nsymbst{2}$, and $\nsymbst{4}$), it applies the symbolic transition $\symbtrans$.
For failed symbolic states ($\nsymbst{3}$ and $\nsymbst{5}$), it applies the
converter $\saconverter$ to convert each of them to the corresponding abstract
state and merges results using the join operator ($\join$).  Then, it utilizes
the abstract semantics via the original view transition
$\viewtrans{\view}{\view'}$ and the converter $\asconverter$ to convert the
result to the corresponding symbolic state.

We could configure when the sealed symbolic execution or the abstract
interpretation is performed based on the definition of the $\checker$ function
and its negation $\neg\checker$.  For example, we could define the $\checker$
that passes only function entry, call, and exit points to perform conversions in
a function level.  However, for the soundness and the termination of the
the abstract interpretation with dynamic shortcut, it should satisfy the
following condition:
\begin{theorem}\label{theorem:shortcut}
  The abstract interpretation with dynamic shortcut is sound and terminates in a
  finite time if its abstract semantics is sound and $\checker$ satisfies the
  following condition:
  \[
    \checker(\symbst) \Rightarrow \symbst \symbtrans^k \excst \wedge
    1 < k \leq N
  \]
  where $N$ is a pre-defined maximum length of the sealed symbolic execution.
\end{theorem}
\inred{We proved Theorem~\ref{theorem:shortcut} using the Kleene's fixed-point
theorem~\cite{kleene} with the finite height of abstract domains. Because of the
page limitation, we omit the proof in this paper and include it in a companion
report~\cite{report}.}
